Review on precise spiking motifs in neurobiological and neuromorphic data

### 2022-10-04 manuscript.tex
importing from manubot

%=================================================================
\documentclass[brainsci,review,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item; 
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}   
\unskip

\begin{table}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}
\unskip

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
			Entry 1		& Data			& Data			& Data\\
			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
\end{table}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a page in landscape format (with table and table footnote).
%\startlandscape
%\begin{table}[H] %% Table in wide page
%\caption{This is a very wide table.\label{tab3}}
%	\begin{tabularx}{\textwidth}{CCCC}
%		\toprule
%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%		\midrule
%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%		\bottomrule
%	\end{tabularx}
%	\begin{adjustwidth}{+\extralength}{0cm}
%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%	\end{adjustwidth}
%\end{table}
%\finishlandscape

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

>>>>

%=================================================================
\documentclass[brainsci,review,submit,pdftex,moreauthors,draft]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

\hypertarget{introduction-importance-of-precise-spike-timings-in-the-brain}{%
\section{Introduction: Importance of precise spike timings in the brain}\label{introduction-importance-of-precise-spike-timings-in-the-brain}}

\hypertarget{ultra-fast-neural-codes-for-ultra-fast-vision}{%
\subsection{Ultra fast neural codes for ultra fast vision}\label{ultra-fast-neural-codes-for-ultra-fast-vision}}

Let us start our review of the state of the art on the role of dynamics in vision by presenting some surprising results that have been obtained in neuroscience. Indeed, Simon Thorpe's group has shown during the last decades numerous examples demonstrating that humans can categorize briefly presented images in a fraction of a second. Its first experiment consisted in asking subjects to categorize images that do or do not contain animals {[}\protect\hyperlink{ref-G3GlupPg}{1}{]}. The results showed that humans were able to perform this task very well (with a success rate of more than 95\%) but above all that a differential activity for the two categories of images could be observed by electroencephalography, showing that this differentiation emerges at a very early latency in neuronal activity. These results have been extended to several species including primates but also extended to different experimental protocols and have shown for example that the response could be extremely fast of the order of 120 ms when the task was to perform a saccade {[}\protect\hyperlink{ref-B9iZWoq3}{2}{]}.

This fast processing also explains the surprising experiments of fast serial detection which consists in presenting a fast succession of different images and to decode via the EEG if the observer can detect for example the presence of an animal. The performances decrease progressively as the frequency of presentation of the images increases. However, it has been shown in the macaque that a significant performance could be maintained with an image presentation time of only 14 ms per image {[}\protect\hyperlink{ref-EqCCVwl2}{3}{]}.

This speed of the visual cortex, although surprising, is quite compatible with the latencies that are recorded at the neuro-physiological level. Indeed, when an image is presented on the retina, the visual information is rapidly propagated to the thalamus and then to the primary visual cortex takes about 55 ms in the macaque {[}\protect\hyperlink{ref-Vu87UU6A}{4}{]}, see Figure {[}\ref{fig:thorpe}{]}. This functioning of visual processing as a forward pass is most prominent in fast processing but can be complemented with feedback loops from the higher areas to the sensory areas {[}\protect\hyperlink{ref-15aMjRHVK}{5}{]}.

\begin{figure}
\hypertarget{fig:thorpe}{%
\centering
\includegraphics[width=1\textwidth,height=\textheight]{https://cdn-learn.adafruit.com/assets/assets/000/086/572/original/circuitpython_Thorpe-FabreThorpe-Neuroscience-Seeking-categories-in-the-brain-monkey-visual-latency-estimate.jpg}
\caption{From {[}\protect\hyperlink{ref-12dWDpY61}{6}{]}}\label{fig:thorpe}
}
\end{figure}

\hypertarget{how-timing-encodes-analogous-profile}{%
\subsection{How timing encodes analogous profile}\label{how-timing-encodes-analogous-profile}}

An important characteristic of neuronal information is that it consists mainly in the transfer of action potentials, or spikes, which consist of brief impulses that propagate along the axons of neurons. These have the particularity of being essentially binary in their amplitude (that is to say, they are prototypical, all or nothing). An important consequence of the speed of processing is that it implies that it is carried out using only very few spikes. Indeed, if we consider that a behavioral response in only 120 ms consists of about ten processing stages following the ``forward'' pathways of the visual system, then this imposes that the processing in a single area is performed with a reduced number of spikes. Neurophysiologists tipcally use firing sequences to characterize the activity of neurons using different statistics on their individual timing {[}\protect\hyperlink{ref-zZUVkTfZ}{7}{]} but also the dependance across neurons {[}\protect\hyperlink{ref-13T2Gne9z}{8}{]}.

Numerous studies demonstrated the importance of precise timing in neural population activity {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}, efficient encoding thanks to the use of spike latencies {[}\protect\hyperlink{ref-1DpSM1kpQ}{10},\protect\hyperlink{ref-IjepoonT}{11}{]} or precise timing in the auditory system {[}\protect\hyperlink{ref-mxPMgOmW}{12},\protect\hyperlink{ref-1ECJhg2GM}{13}{]}. All these findings, and more {[}\protect\hyperlink{ref-WZsEyOlK}{14}{]}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in the spike train. Note also that depending on how an in vitro cell is driven, this may influence the reliability of spike timing in neocortical neurons {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]}.

\begin{figure}
\hypertarget{fig:roc}{%
\centering
\includegraphics{images/roc.jpg}
\caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron. From {[}\protect\hyperlink{ref-YX5wqWH4}{16}{]}}\label{fig:roc}
}
\end{figure}

At the level of dynamic processing of visual information, it has been shown that an encoding of the values of luminance imminence in the image instead of the retina {[}\protect\hyperlink{ref-IjepoonT}{11}{]}. Notably one can appreciate in figure 1 that the response of ganglion cells to visual gratings that are flashed onto the retina. The authors showed that the neuronal response could be encoded in the latency of the response and not only in the frequency of discharge as is often assumed. In figure 4 of the same article, these results are extended to natural images and show a qualitatively similar behavior. The conclusion of the authors is that the discharge latency of the neurons allows to encode spatial characteristics of the image. First-spike latency codes are a feasible mechanism for information transfer even when biologically plausible estimates of stimulus onset are taken into account {[}\protect\hyperlink{ref-1H1Omh73g}{17}{]}. Interestingly, such fine latency effect can be at the origin of somme visual illusions, for instance the illusion of false colors in the \href{https://michaelbach.de/ot/col-Benham/index.html}{Benham Top} based based on center-surround interactions in the parvocellular pathway {[}\protect\hyperlink{ref-UlbjMU0W}{18}{]}.

Similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area {[}\protect\hyperlink{ref-513xqzmL}{19}{]}. Many models have used these properties in temporal coding to build fast image categorization networks. These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization {[}\protect\hyperlink{ref-1DpSM1kpQ}{10}{]}. This work has been extended to include unsupervised learning capabilities and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes {[}\protect\hyperlink{ref-7K80jcx9}{20},\protect\hyperlink{ref-fRudGmR1}{21}{]}. This type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Cameras which, instead of using a basic frame-based representation, uses a representation similar to the one we have just described and which consists in representing the image by events {[}\protect\hyperlink{ref-Oet1bgdb}{22}{]}. This type of modeling often uses the classical architecture of image categorization developed in deep learning while adapting it to the specificity of the event-based representation {[}\protect\hyperlink{ref-15cR83gqd}{23}{]}. Note also that timing is not entirely sensorial or internal but in {[}\protect\hyperlink{ref-1E8gkynCm}{24}{]}, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines. Timing, at least in animals, may thus be fundamentally embodied and situated.'' Altogether, these results show that ``that like other senses, vision relies heavily on temporal strategies and temporal neural codes to extract and represent spatial information'' {[}\protect\hyperlink{ref-QK8RfvTZ}{25}{]}.

\hypertarget{from-synfire-chains}{%
\subsection{From synfire chains\ldots{}}\label{from-synfire-chains}}

The analysis of generic raster plots reveals particular traits that hint at the role of a precise timing. For instance, the firing rate is highly irregular {[}\protect\hyperlink{ref-sYHyiGsP}{26}{]}, which makes it inconsistent with temporal integration of random EPSPs. It has also been observed that the response of a neuron in a cortical slice to a current step could be highly non-reproducible, where the first spike is aligned but the subsequent spike times tend to diffuse for independent repetitions of the stimulation {[}\protect\hyperlink{ref-XsHCA4IZ}{15}{]} (see Figure {[}\ref{fig:mainen}{]}). When driven by a highly dynamic \emph{frozen} noise, the output spikes are highly reproducible. This is consistent with the differential role of different stimulus frequencies on the reliability of spike timing reported in {[}\protect\hyperlink{ref-1AbvCTG03}{27}{]} : ``we found that, as expected given the resistive and capacitive properties of cortical neurons, low frequencies have a larger effect on the membrane potential of cortical neurons than do higher frequencies. However, increasing the amount of gamma range fluctuations in a stimulus leads to more precise timing of action potentials.''

\begin{figure}
\hypertarget{fig:mainen}{%
\centering
\includegraphics{https://camo.githubusercontent.com/77f4a48a0fe8f98ffab3c1f21dca1e74286b14d47c17120ee9627f3351fcd5e7/687474703a2f2f692e737461636b2e696d6775722e636f6d2f69786e727a2e706e67}
\caption{The timing of the spikes produced following the repetition of a step stimuli is less reproducible than that to a noisy stimulus. While this seems paradoxical at first sight, this is due to the use of the same \emph{frozen} nois at each repetition and highlight the highly reproducible pattern of spikes when it is driven by a highly dynamic input. See this \href{https://github.com/laurentperrinet/2022_UE-neurosciences-computationnelles/blob/master/C_MainenSejnowski1995.ipynb}{notebook} for a replication of this result using a simple LIF model.}\label{fig:mainen}
}
\end{figure}

In his book ``Corticonics'' {[}\protect\hyperlink{ref-19MS6Bbwp}{28}{]}, Moshe Abeles queried if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns. The book gradually leads the reader from the macroscopic cortical anatomy and standard electrophysiological properties of single neurons to neural network models. While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code {[}\protect\hyperlink{ref-lT1Ato7P}{29}{]}. The book then demonstrates that this could take the form of ``synfire chains,'' that is synchronous activity on subsets of neurons which could be propagated in a stable fashion. Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons {[}\protect\hyperlink{ref-R2E2wRFg}{32}{]}.

It was shown that a simple model may allow the propagation of such synfire chains {[}\protect\hyperlink{ref-xbqYi2PD}{33}{]}. This model considers the dynamics of leaky integrate-of-fire neurons in different groups of similar size. Each neuron of one group is connected by an excitatory synapse to the next. When a pulse is elicited in the first group, this may generate a spike in the next group. Depending on the weight value, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse). Recursively applying this to a sequence of groups generates either a synfire propagation or not. A simple simulation is shown in Figure {[}\ref{fig:diesman}{]}. A crucial aspect of this emergence is in particular the balance between excitation and inhibition {[}\protect\hyperlink{ref-DL2NpHxr}{34}{]}. This was for instance modeled by feed-forward inhibition which is a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex {[}\protect\hyperlink{ref-mszJ8wdv}{35}{]}.

\begin{figure}
\hypertarget{fig:diesman}{%
\centering
\includegraphics{https://brian2.readthedocs.io/en/stable/_images/frompapers.Diesmann_et_al_1999.1.png}
\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}. The model consists of 10 groups of 100 neurons each. A pulse with a given jitter is generated in the first group which here generates a pulse after a certain processing delay in the second group with a lower jitter. This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
}
\end{figure}

Attempts have been made to detect such synfire chains in neurobiological data {[}\protect\hyperlink{ref-qVu1OkoS}{36}{]}: ``The sensitivity is high enough to detect synfire chain activity in simultaneous single-unit recordings of 100 to 200 neurons from such data, enabling application to experimental data in the near future.'' Further models have shown that such synfire chains could be embedded in topographies {[}\protect\hyperlink{ref-xBZqHOuN}{37}{]} or using conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation {[}\protect\hyperlink{ref-p7xk5w4f}{38}{]}. In particular, this was implemented using the pyNN language {[}\protect\hyperlink{ref-33SRV9MB}{39}{]} both in CPU-based and neuromorphic hardware {[}\protect\hyperlink{ref-DiIieN8s}{40}{]} (see Figure {[}\ref{fig:pynn}{]}).

\begin{figure}
\hypertarget{fig:pynn}{%
\centering
\includegraphics{https://www.frontiersin.org/files/Articles/39056/fnins-07-00011-r2/image_m/fnins-07-00011-g004.jpg}
\caption{A pyNN implementation of a Synfire chain with feedforward inhibition. The background is only utilized in the original model, where it is implemented as random Gaussian current. For the presented hardware implementation it has been omitted due to network size constraints. As compensation for missing background stimuli, the resting potential was increased to ensure a comparable excitability of the neurons. The state space gives the duration of synfire chains as a function of the protocol's parameters and compares (C) CPU to (D) neuromorphic implementations.}\label{fig:pynn}
}
\end{figure}

Note also that synchronicity may explain some unintuitive results. Indeed it has been shown that thalamocortical synapses are relatively weak compared to the amount of intra-cortical activity. However, this pathway is sufficient to drive the cortex as this input is more often synchronously active {[}\protect\hyperlink{ref-18ZU9zP9l}{41}{]}.

Some experimental results show the emergence of synchrony, for instance in motor cortical function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}. Interestingly, authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task {[}\protect\hyperlink{ref-1HJjmmSpm}{43}{]} and showed some tuning to to movement direction and reaction time {[}\protect\hyperlink{ref-FZ9X31Qa}{44}{]}. It is important to note that synchronous events tended to lock to LFP beta waves {[}\protect\hyperlink{ref-Ejy1sAyh}{45}{]}, and was extended to larger assemblies {[}\protect\hyperlink{ref-Glu4NlLC}{47}{]} using statistical methods (see {[}\protect\hyperlink{ref-sec:detection}{\textbf{sec:detection?}}{]}).

In {[}\protect\hyperlink{ref-3GlNNxH5}{48}{]}, the author proposes a simple spike-based computational framework, based on the idea that stimulus-induced synchrony can be used to extract sensory invariants (for example, the location of a sound source), which is a difficult task for classical neural networks. It relies on the simple remark that a series of repeated coincidences is in itself an invariant. Many aspects of perception rely on extracting invariant features, such as the spatial location of a time-varying sound, the identity of an odor with fluctuating intensity, the pitch of a musical note.

\hypertarget{to-traveling-waves}{%
\subsubsection{.. to traveling waves \ldots{}}\label{to-traveling-waves}}

TODO : {[}\protect\hyperlink{ref-zNDCo9Hy}{49}{]}

Statistical dependencies in the responses of sensory neurons govern both the amount of stimulus information conveyed and the means by which downstream neurons can extract it. This is put in evidence by analysing the functional significance of correlated firing in a complete population of macaque parasol retinal ganglion cells using a model of multi-neuron spike responses {[}\protect\hyperlink{ref-Piao2Wzf}{50}{]}. This show precise spatio-temporal differences in this recurrently connected assembly. The different aspects of information in the data are evaluated by a decoding strategy highlighting the role of correlations. A similar dataset used in {[}\protect\hyperlink{ref-2AsQOOsE}{52}{]} is available from M. Berry's lab {[}\protect\hyperlink{ref-1FU0Y7Erc}{53}{]}.

Propagating waves occur in many excitable media and were recently found in neural systems from retina to neocortex. While propagating waves are clearly present under anaesthesia, whether they also appear during awake and conscious states remains unclear. One possibility is that these waves are systematically missed in trial-averaged data, due to variability. A recent work {[}\protect\hyperlink{ref-hGYLURVS}{54}{]} presents a method for detecting propagating waves in noisy multichannel recordings. Applying this method to single-trial voltage-sensitive dye imaging data, they show that the stimulus-evoked population response in primary visual cortex of the awake monkey propagates as a travelling wave, with consistent dynamics across trials. A network model suggests that this reliability is the hallmark of the horizontal fibre network of superficial cortical layers. Propagating waves with similar properties occur independently in secondary visual cortex, but maintain precise phase relations with the waves in primary visual cortex. These results show that, in response to a visual stimulus, propagating waves are systematically evoked in several visual areas, generating a consistent spatiotemporal frame for further neuronal interactions.

\begin{figure}
\hypertarget{fig:davis}{%
\centering
\includegraphics[width=1\textwidth,height=\textheight]{https://media.springernature.com/lw685/springer-static/image/art\%3A10.1038\%2Fs41467-021-26175-1/MediaObjects/41467_2021_26175_Fig1_HTML.png}
\caption{a Spike raster for repeated presentations (N = 40) of a high-contrast (10\% Michelson contrast) drifting Gabor recorded from area MT of a fixating marmoset (stimulus-onset, red line; mean response, blue line). A single-trial LFP trace is plotted in gray, and the average LFP response is plotted in black. The relative power between baseline (−200 ms to stimulus-onset) and evoked fluctuations (stimulus-onset + 50--250 ms) significantly favored the evoked response (right panel; N = 110 trials; median = 1.89 dB, p = 0.000019, two-tailed Wilcoxon's ranked-sum test). b Same as in (a), but for a low contrast stimulus (\textless2\% Michelson contrast). The relative power between baseline and evoked LFP fluctuations was not statistically different from parity (median = 1.23 dB, p = 0.087 two-tailed Wilcoxon's ranked-sum test). c An example of spontaneous LFP fluctuations structured as a traveling wave recorded from a spatially distributed multielectrode array in marmoset area MT. d Histogram of spontaneous spike probability as a function of the generalized phase of the LFP during fixation. e The average evoked response to low contrast stimuli was stronger when a more excitable phase (±π rad) of a spontaneous traveling wave aligned with the retinotopic location of the target (aligned, green line) as compared to when a less excitable phase (0 rad) was aligned (unaligned, purple line; N = 43 wave and non-wave trials; shaded region SEM; p = 0.0000015 two-tailed Wilcoxon's rank-sum test). Data for panels c--e modified from Davis et al.17 with permission.}\label{fig:davis}
}
\end{figure}

More recently Advanced recording techniques have enabled the identification of travelling waves of neuronal activity in different areas of the cortex {[}\protect\hyperlink{ref-vsKvZJ82}{55}{]}. Authors review these findings, consider the mechanisms by which travelling waves are generated and evaluate their possible roles in cortical function. In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states {[}\protect\hyperlink{ref-BNK6T5Uk}{9}{]}. Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise. However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity. The mechanisms underlying these waves are unknown. Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings. Here, authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales. They find that spontaneous traveling waves are a general property of these networks. The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave. Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states. Further, by modulating local network state, they can shape responses to incoming inputs as observed in vivo.

{[}\protect\hyperlink{ref-17EU8fPBv}{56}{]} present ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.

\hypertarget{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}{%
\subsubsection{\texorpdfstring{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}\protect\hyperlink{ref-wVZpQvSk}{42}{]}}{Spike Synchronization and Rate Modulation Differentially Involved in Motor Cortical Function {[}42{]}}}\label{spike-synchronization-and-rate-modulation-differentially-involved-in-motor-cortical-function-doi10.1126science.278.5345.1950}}

It is now commonly accepted that planning and execution of movements are based on distributed processing by neuronal populations in motor cortical areas {[}\protect\hyperlink{ref-yLVM9IXi}{57}{]}. It is less clear, though, how these populations organize dynamically to cope with the momentary computational demands. Simultaneously recorded activities of neurons in the primary motor cortex of monkeys during performance of a delayed-pointing task exhibited context-dependent, rapid changes in the patterns of coincident action potentials. Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.

{[}\protect\hyperlink{ref-OAorhMRa}{58}{]}

\hypertarget{todo-and-to-synfire-braids-and-polychronous-groups}{%
\subsubsection{TODO \ldots{} and to synfire braids and polychronous groups}\label{todo-and-to-synfire-braids-and-polychronous-groups}}

\begin{itemize}
\tightlist
\item
  {[}\protect\hyperlink{ref-HjDb1gnq}{59}{]} : from synfire chains to Synfire braids
\end{itemize}

In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays {[}\protect\hyperlink{ref-KS4oxnst}{60},\protect\hyperlink{ref-ga3ZTFC7}{61}{]}. Notably, Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]} introduced the notion of polychronous groups (PGs) as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, spiking delays. This representation has a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs.

sparse in time and space {[}2{]} AL Barth and JF Poulet Trends in Neurosciences 35.6 (2012), pp.~345-355. {[}3{]} CC Petersen and S Crochet, Neuron 78.1 (2013), pp.~28-48.

\begin{itemize}
\item
  recent theories of binding by synchrony : Fries 2005 trends cog neuro with spikes arriving at peak susceptibility (top of a cycle) or down, van Rullen, Laura Dugué
\item
  cell assembly hypothesis: neurons coordinate their activity through the formation and repetitive co-activation of groups. https://link-springer-com.insb.bib.cnrs.fr/article/10.1007/s10827-021-00801-9 Hebb DO. The organisation of behaviour: a neuropsychological theory. New York: Science Editions; 1949.
\item
  A notable exception is the polychronization model of Izhikevich {[}\protect\hyperlink{ref-SM9G0xBK}{62}{]}, which combined the construction of a random recurrent model of spiking neurons including such delays and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule. In this model, raster plot analysis showed repeated activation of Polychronous Groups (PGs), i.e., specific spike patterns with a specific sequence of activations.
\end{itemize}

\hypertarget{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}{%
\subsection{Are there precise heterosynaptic spiking motifs in the brain?}\label{are-there-precise-heterosynaptic-spiking-motifs-in-the-brain}}

Currently, there is a consensus for rate-based coding models in computational and biological neuroscience. Nevertheless, there is a substantial literature in neurobiology indicating that brain dynamics often organize into stereotyped sequences (like synfire chains {[}\protect\hyperlink{ref-Mf6h1KNY}{63}{]}, packets {[}\protect\hyperlink{ref-2GovoaYw}{64}{]} or hippocampal sequences {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and on the role of such precise spike timing in downstream information transfer and coding {[}\protect\hyperlink{ref-19fAM6hcv}{67},\protect\hyperlink{ref-13dQ0R4UM}{68},\protect\hyperlink{ref-MfmcMsqA}{69}{]}. This is for instance relevant in sensory pathways in vision {[}\protect\hyperlink{ref-1EHZgcKyL}{70}{]}, audition {[}\protect\hyperlink{ref-FlKJw2VL}{71}{]}, olfaction {[}\protect\hyperlink{ref-AhSRXBJ2}{72}{]}{]} or touch {[}\protect\hyperlink{ref-gDinei4a}{73}{]}. In particular, one theoretical viewpoint considers synfire braids (Bienenstock, 1995), where a precise sequential motif of spikes will synchronize as it reaches the soma of a neuron for which synaptic delays are adequately tuned. In particular, computational modeling shows that at the scale of neurons, an efficient neural code can emerge where spike times are organized in prototypical, precise temporal motifs (Izhikevich, 2006) which he defined as polychronous groups.

Stereotyped sequences of neuronal activation have been particularly well described in the adult hippocampus and related to its function in mental travel in time and space {[}\protect\hyperlink{ref-13dQ0R4UM}{68}{]}. These sequences can be internally generated {[}\protect\hyperlink{ref-SVfasyCp}{65},\protect\hyperlink{ref-PUNGVou}{66}{]} and are formed by the chained activation of orthogonal assemblies, themselves organized as sequence packets (Malvache et al., 2016). Thus, hippocampal sequences are formed by the ordered activation of smaller sequence motifs. They are stereotyped and robust, since neurons can be activated in the same order across days (see Figure {[}\ref{fig:haimerl}{]} from {[}\protect\hyperlink{ref-qI2HUEyy}{74}{]} below). As a consequence, hippocampal sequences may rely on an internally hardwired structure and form the functional building blocks for encoding, storing and retrieving experience.

\begin{figure}
\hypertarget{fig:haimerl}{%
\centering
\includegraphics{https://www.pnas.org/cms/10.1073/pnas.1718518116/asset/ab9a48f6-b28f-40c1-978e-6f6dbec39b98/assets/graphic/pnas.1718518116fig01.jpeg}
\caption{Mixed distance and duration representation in CA1. (A) Calcium fluorescence (heatmap) of CA1 neurons participating to run sequences in consecutive imaging sessions. Cells have been selected and ordered with respect to their activity in the first imaging session (Top). The black line represents speed of the mouse.}\label{fig:haimerl}
}
\end{figure}

TODO: read {[}\protect\hyperlink{ref-MfmcMsqA}{69}{]} Luczak A, McNaughton BL, Harris KD. Packet-based communication in the cortex. Nat Rev Neurosci. 2015;16(12):745--55.
